### 2012-04-09 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: Corrections.

The following text is an example of inference of a more complex query processing. Consider the following part of a query.
\begin{verbatim}
declare function local:getB($id as xs:string) as element() {
  //A[@id = $id]/B
}
... local:getBs("id") > 10 ...
\end{verbatim}
The query consists of a function declaration and an arithmetic comparison. The comparison compares the result of the function call and literal value \texttt{10}. For the type of \texttt{10} is \texttt{xs:integer} the type of the function call has to be convertible to \texttt{xs:double}. Thus, it has to be a numeric type. The function returns a path expression typed as \texttt{element()}. That means the function returns one element. In the path expression, the argument \texttt{\$id} can be substituted by the real value specified in the call. Thus, the return expression is \texttt{//A[@id = "id"]/B}. Therefore, we can infer that element \texttt{B} in element \texttt{A} with attribute \texttt{id} equaling string value \texttt{"id"} is of some numeric type. And, since the function is parametrized, there is a notion that this statement may be correct also for other elements \texttt{A} and \texttt{B}.

>>>>

The following text is an example of inference of a more complex query processing. Consider the following part of a query.
\begin{verbatim}
declare function local:getB($id as xs:string) as element() {
  //A[@id = $id]/B
}
... local:getBs("id") > 10 ...
\end{verbatim}
The query consists of a function declaration and an arithmetic comparison. The comparison compares the result of the function call and literal value \texttt{10}. For the type of \texttt{10} is \texttt{xs:integer} the type of the function call has to be convertible to \texttt{xs:double}. Thus, it has to be a numeric type. The function returns a path expression typed as \texttt{element()}. That means the function returns one element. In the path expression, the argument \texttt{\$id} can be substituted by the real value specified in the call. Thus, the return expression is \texttt{//A[@id = "id"]/B}. Therefore, we can infer that element \texttt{B} in element \texttt{A} with attribute \texttt{id} equalling string value \texttt{"id"} is of some numeric type. And, since the function is parametrized, there is a notion that this statement may be correct also for other elements \texttt{A} and \texttt{B}.

### 2012-04-09 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections.

Without loss of generality, we will focus on locally defined functions. We will not determine types of functions defined in other modules since the principle is similar but it is needed to look up the definitions in syntax trees of other queries. Return types of built-in functions are fixed and, thus, there is no need to analyse them. Also, since a determination of a prefix for built-in functions is a technical issue, we assume that built-in functions are either prefixed by \texttt{fn} or not prefixed.

@@

\subsection{Determination of Global Variable Types}
A similar approach as in the case of functions can be applied to determine types of global variables. Alike the functions, the global variables are defined in the prolog section. A type of a variable can be explicitly specified in its definition, for instance \texttt{declare variable \$x as xs:byte := 12;}. If it is not, it may often be deducible from the binding expression. Again, we do not analyse external variables for the same reason as in case of external functions.

@@

If a currently processed expression is a function call of an user-defined function and at least one of its real arguments is an instance of PathType, get the function declaration node. In the function body, replace the formal arguments by the real arguments and set them as the local variables, so the proposed algorithms can be applied to process the function's body. Analyse the types of expressions in the body with the new information on the types of arguments by incorporating of the proposed algorithms and replace the function's return type by the one currently determined.

>>>>

Without loss of generality, we will focus on locally defined functions. We will not determine types of functions defined in other modules since the principle is similar but it is needed to look up the definitions in syntax trees of other queries. Return types of built-in functions are fixed and, thus, there is no need to analyze them. Also, since a determination of a prefix for built-in functions is a technical issue, we assume that built-in functions are either prefixed by \texttt{fn} or not prefixed.

@@

\subsection{Determination of Global Variable Types}
A similar approach as in the case of functions can be applied to determine types of global variables. Alike the functions, the global variables are defined in the prolog section. A type of a variable can be explicitly specified in its definition, for instance \texttt{declare variable \$x as xs:byte := 12;}. If it is not, it may often be deducible from the binding expression. Again, we do not analyze external variables for the same reason as in case of external functions.

@@

If a currently processed expression is a function call of an user-defined function and at least one of its real arguments is an instance of PathType, get the function declaration node. In the function body, replace the formal arguments by the real arguments and set them as the local variables, so the proposed algorithms can be applied to process the function's body. Analyze the types of expressions in the body with the new information on the types of arguments by incorporating of the proposed algorithms and replace the function's return type by the one currently determined.

### 2012-04-09 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: Corrections and final(?) additions.
Master thesis: Corrections.

\begin{figure}
\begin{verbatim}
<bookstore>
  <book id="b1">
    <title>
      Compilers: Principles, Techniques, and Tools
      (2nd Edition)
    </title>
    <authors>
      <author>Alfred V. Aho</author>
      <author>Monica S. Lam</author>
      <author>Ravi Sethi</author>
      <author>Jeffrey D. Ullman</author>
    </authors>
  </book>
  <book id="b2">
    <title>XQuery</title>
    <authors>
      <author>Priscilla Walmsley</author>
    </authors>
    <isbn>0596006349</isbn>
  </book>
</bookstore>
\end{verbatim}
\caption{A simple XML example.}
\label{FIG_a_simple_xml_example}
\end{figure}

@@

The dafault value is either a literal value or one of the following specifiers.

@@

<!ELEMENT isbn #PCDATA>
\end{verbatim}
\caption{A simple DTD example.}
\label{FIG_a_simple_dtd_example}
\end{figure}

@@

The XSD language consists of many more constructs we do not mention such as substitution groups, type extensions, integrity constranints and other.

@@

\begin{figure}
\begin{alltt}
<xs:element name="book">
  <xs:complexType>
    <xs:sequence>
      <xs:element name="title" type="xs:string"/>
      <xs:element name="authors>
        <xs:complexType>
          <xs:element name="author" type="xs:string"
           maxOccurs="unbound"/>
        </xs:complexType>
      </xs:element>
      <xs:element name="isbn" type="xs:string" minOccurs="0"/>
    </xs:sequence>
    <xs:attribute name="id" type="xs:string" use="required"/>
  </xs:complexContent>
</xs:element>
\end{alltt}
\caption{A simple XSD example.}
\label{FIG_a_simple_xsd_example}
\end{figure}

@@

\subsection{Abbreviations}
Abbreviations for several most widely used constructs are defined as follows.
\begin{itemize}
\item \texttt{$P$/$identifier$} stands for \texttt{$P$/child::$identifier$} (child is a default axis)
\item \texttt{$P$/@$identifier$} stands for \texttt{$P$/atribute::$identifier$}
\item \texttt{$P$/../$identifier$} stands for \texttt{$P$/parent::*/$identifier$}
\item \texttt{$P$//$identifier$} stands for \texttt{$P$/descendant-or-self::node()/$identifier$} 
\end{itemize}

@@

\subsection{FLWOR Expressions}
A basic construct of the XQuery language is $FLWOR$. It is an abbrevation of its five clauses: for, let, where, order by, return. For clause is \\
\texttt{for $var$ in $expr$} \\
Expression $expr$ is evaluated and its result is a sequence. Items of the sequence are iteratively assigned to $var$ variable, which is valid also in the following clauses.

>>>>

\begin{figure}
\begin{verbatim}
<bookstore>
  <book id="b1">
    <title>
      Compilers: Principles, Techniques, and Tools
      (2nd Edition)
    </title>
    <authors>
      <author>Alfred V. Aho</author>
      <author>Monica S. Lam</author>
      <author>Ravi Sethi</author>
      <author>Jeffrey D. Ullman</author>
    </authors>
  </book>
  <book id="b2">
    <title>XQuery</title>
    <authors>
      <author>Priscilla Walmsley</author>
    </authors>
    <isbn>0596006349</isbn>
  </book>
</bookstore>
\end{verbatim}
\caption{A simple XML example}
\label{FIG_a_simple_xml_example}
\end{figure}

@@

The default value is either a literal value or one of the following specifiers.

@@

<!ELEMENT isbn #PCDATA>
\end{verbatim}
\caption{A simple DTD example}
\label{FIG_a_simple_dtd_example}
\end{figure}

@@

The XSD language consists of many more constructs we do not mention such as substitution groups, type extensions, integrity constraints and other.

@@

\begin{figure}
\begin{alltt}
<xs:element name="book">
  <xs:complexType>
    <xs:sequence>
      <xs:element name="title" type="xs:string"/>
      <xs:element name="authors>
        <xs:complexType>
          <xs:element name="author" type="xs:string"
           maxOccurs="unbound"/>
        </xs:complexType>
      </xs:element>
      <xs:element name="isbn" type="xs:string" minOccurs="0"/>
    </xs:sequence>
    <xs:attribute name="id" type="xs:string" use="required"/>
  </xs:complexContent>
</xs:element>
\end{alltt}
\caption{A simple XSD example}
\label{FIG_a_simple_xsd_example}
\end{figure}

@@

\subsection{Abbreviations}
Abbreviations for several most widely used constructs are defined as follows.
\begin{itemize}
\item \texttt{$P$/$identifier$} stands for \texttt{$P$/child::$identifier$} (child is a default axis)
\item \texttt{$P$/@$identifier$} stands for \texttt{$P$/attribute::$identifier$}
\item \texttt{$P$/../$identifier$} stands for \texttt{$P$/parent::*/$identifier$}
\item \texttt{$P$//$identifier$} stands for \texttt{$P$/descendant-or-self::node()/$identifier$} 
\end{itemize}

@@

\subsection{FLWOR Expressions}
A basic construct of the XQuery language is $FLWOR$. It is an abbreviation of its five clauses: for, let, where, order by, return. For clause is \\
\texttt{for $var$ in $expr$} \\
Expression $expr$ is evaluated and its result is a sequence. Items of the sequence are iteratively assigned to $var$ variable, which is valid also in the following clauses.
